---
title: "Assignment 1"
author: "Natalie Benoy"
date: "1/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Data & Packages
```{r}
library(tidyverse)
usedcars <- read_csv("usedcars.csv")
houseprices <- read_csv("1b_HousePrices.csv")
toyota <- read_csv("1c_ToyotaCorolla.csv")
```

```{r}
# turn character variables into factors
usedcars <- usedcars %>%
  mutate(model = factor(model),
         color = factor(color),
         transmission = factor(transmission))

glimpse(usedcars)

# get observations of cars with prices > 18000
usedcars$price > 18000
usedcars[usedcars$price > 18000,]

# retrieve observations of cars with price > 18000 and color = yellow
usedcars[usedcars$price > 18000 & usedcars$color == "Yellow",]

# interquartile range of price
IQR(usedcars$price)

# percentiles of price at 20% increments
quantile(usedcars$price, c(0, 0.2, 0.4, 0.6, 0.8, 1))

# histogram of price
ggplot(usedcars, aes(price)) + geom_histogram()

# boxplot of price
ggplot(usedcars, aes(y = price)) + geom_boxplot()

# boxplot of price by model
ggplot(usedcars, aes(x = model, y = price)) + geom_boxplot()

# use crosstable
library(gmodels)
CrossTable(usedcars$transmission, usedcars$model)
```

## Assignment 1A

# Answer 1
```{r}
student_name <- c("Bob Dole", "Joe Schmo", "Betty White")
gpa <- c(3.2, 3.5, 3.7)
gender <- c("M", "M", "F")
grade <- c("junior", "sophomore", "senior")
```

# Answer 2: M, junior, sophomore
```{r}
gender[2]
grade[1:2]
```

# Answer 3
```{r}
gender <- factor(c("M", "M", "F"), levels = c("F", "M"))
grade <- factor(c("junior", "sophomore", "senior"), levels = c("freshman", "sophomore", "junior", "senior"), ordered = T)
```

# Answer 4: Rows: 3, Columns: 4
```{r}
dataframe <- data.frame(student_name, gpa, gender, grade, stringsAsFactors = F)

dataframe
```

# Answer 5
```{r}
dataframe[,2]

dataframe[c("student_name", "grade")]
```

# Answer 6: 3.2
```{r}
dataframe[1,2]
```

# Answer 7
```{r}
dataframe[c(2,3),c(3,4)]
```

# Answer 8
```{r}
dataframe[,2]
dataframe[3,]
dataframe[,]
```

# Answer 9
```{r}
M <- matrix(c("a", "b", "c", "d", "e", "f", "g", "h"), nrow = 4)
M

N <-  matrix(c("a", "b", "c", "d", "e", "f", "g", "h", "i"), nrow = 3)
N
```

# Answer 10: b, f
```{r}
M[2,1]
N[3,2]
```

# Answer 11
```{r}
M[2,]
N[,3]
```

## Assignment 1B

# Answer 1: 128 houses
```{r}
houseprices <- read_csv("1b_HousePrices.csv")
str(houseprices)
```

# Answer 2: Max house price: 211200, median offers: 3, mean SqFt: 2001
```{r}
summary(houseprices)
```

# Answer 3: mean bedrooms: 3.023, median bedrooms: 3
```{r}
mean(houseprices$Bedrooms)
median(houseprices$Bedrooms)
```

# Answer 4: range: 69100-211200, diff: 142100
```{r}
range(houseprices$Price)
diff(range(houseprices$Price))
```

# Answer 5: 36925
```{r}
IQR(houseprices$Price)
```

# Answer 6:
```{r}
quantile(houseprices$Price, c(0, 0.2, 0.4, 0.6, 0.8, 1))
```

# Answer 7: middle line represents the median
```{r}
ggplot(houseprices, aes(y = Price)) + geom_boxplot()

ggplot(houseprices, aes(y = SqFt)) + geom_boxplot()
```

# Answer 8: Highest frequency price range: 115000-120000 Highest frequency offer range: 3
```{r}
ggplot(houseprices, aes(x = Price)) + geom_histogram(binwidth = 5000)

ggplot(houseprices, aes(x = Offers)) + geom_histogram()
```

# Answer 9: variance: 721930821, sd: 26868.77
```{r}
var(houseprices$Price)
sd(houseprices$Price)
```

# Answer 10: 42 houses built with bricks, 39 houses on West side
```{r}
table(houseprices$Brick)
table(houseprices$Neighborhood)
```

# Answer 11: 35.16% of houses are on East side
```{r}
neighborhood_table <- table(houseprices$Neighborhood)
prop.table(neighborhood_table)
```

# Answer 12: 32.81% of houses are brick
```{r}
brick_table <- table(houseprices$Brick)
prop.table(brick_table)
```

# Answer 13: All other things equal, as SqFt increases, Price increases
```{r}
ggplot(houseprices, aes(x = SqFt, y = Price)) + geom_point()
```

# Answer 14: houses built w/brick in north neighborhood: 7
```{r}
library(gmodels)
CrossTable(houseprices$Brick, houseprices$Neighborhood)
```

## Assignment 1C

# Answer 1: 1436 cars in the dataset
```{r}
toyota <- read_csv("1c_ToyotaCorolla.csv")
str(toyota)
```

# Answer 2: Max car price: 32500, median car age: 61, mean weight: 1072, mean KM: 68533
```{r}
summary(toyota)
```

# Answer 3: mean HP: 101.50 median HP: 110
```{r}
mean(toyota$HP)
median(toyota$HP)
```

# Answer 4: range: 4350-32500, diff: 28150
```{r}
range(toyota$Price)
diff(range(toyota$Price))
```

# Answer 5: IQR: 3500
```{r}
IQR(toyota$Price)
```

# Answer 6:
```{r}
quantile(toyota$Price, c(0, 0.25, 0.5, 0.75, 1))
```

# Answer 7: Bottom line stands for the lowest point 1.5 * IQR below the lower quartile (anything below that point is considered an outlier on the low end)
```{r}
ggplot(toyota, aes(y = Age)) + geom_boxplot()
ggplot(toyota, aes(y = KM)) + geom_boxplot()
```

# Answer 8: Highest frequency price range: 8500-9500, Highest freqency weight range: 1040-1060
```{r}
ggplot(toyota, aes(x = Price)) + geom_histogram(binwidth = 1000)
ggplot(toyota, aes(x = Weight)) + geom_histogram(binwidth = 20)
```

# Answer 9: Variance: 13154872, Sd: 3626.97
```{r}
var(toyota$Price)
sd(toyota$Price)
```

# Answer 10: 969 cars are metallic, 674 cars have 5 doors, 17 cars use CNG as fuel type
```{r}
table(toyota$MetColor)
table(toyota$Doors)
table(toyota$FuelType)
```

# Answer 11: 0.10793872 proportion of cars use Diesel
```{r}
FuelTable <- table(toyota$FuelType)
prop.table(FuelTable)
```

# Answer 12:10.79% of cars use Diesel
```{r}
FuelTable <- table(toyota$FuelType)
prop.table(FuelTable)
```

# Answer 13: As KM increases, Price decreases non-linearly toward a lower limit > 0
```{r}
ggplot(toyota, aes(x = KM, y = Price)) + geom_point()
```

# Answer 14: Cars w/automatic transmission, diesel: 0. Cars w/automatic transmission & 5 doors: 30
```{r}
CrossTable(toyota$Automatic, toyota$FuelType)
CrossTable(toyota$Automatic, toyota$Doors)
```
